\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-01-23 papers/jss/article.Rnw
Add two minor text improvements suggested by Tim Hesterberg to sections 5 and 6.
The first sentence of each of the last two paragraphs of the summary
was saying the same thing -- merge the second duplicate sentence into
the first, which then leaves the third paragraph of the summary as
just a single sentence.  Add it to the end of the second sentence and
we now have I think a slightly crisper summary in just two paragraphs.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@832 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Table~\ref{table-get-types} details the correspondence between the
field type and the type of data that is retrieved by \verb|$| and \verb|[[|
extractors.  Three types in particular need further attention due to
specific differences in the \proglang{R} language.

@@

\Sexpr{m} data sets can be converted to Protocol Buffers
(\Sexpr{format(100*m/n,digits=1)}\%) without loss of information. Upon closer
inspection, all other datasets are objects of class \texttt{nfnGroupedData}.
This class represents a special type of data frame that has some additional 
attributes (such as a \emph{formula} object) used by the \pkg{nlme} package.
Because formulas are \proglang{R} \emph{language} objects, they have little meaning to
other systems, and are not supported by the \texttt{rexp.proto} descriptor.
When \texttt{serialize\_pb} is used on objects of this class, it will serialize
the data frame and all attributes, except for the formula.

@@

\section{Summary}  % DE Simpler title
\label{sec:summary}
% TODO(mstokely): Get cibona approval for these two sentences before
% publishing
Over the past decade, many formats for interoperable
data exchange have become available, each with their unique features,
strengths and weaknesses.  
Text based formats such as CSV and JSON are easy to use, and will likely 
remain popular among statisticians for many years to come. However, in the 
context of increasingly complex analysis stacks and applications involving 
distributed computing as well as mixed language analysis pipelines, choosing a more 
sophisticated data interchange format may reap considerable benefits. 
The Protocol Buffers protocol and library offers a unique combination of features, performance,
maturity, and forward-compatibility that seems particulary well suited for data-driven 
applications and numerical computing.

%% DE Re-ordering so that we end on RProtoBuf
The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
extends the \proglang{R} system with the ability to create, read, and write Protocol
Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
for the past three years by statisticians, analysts, and software engineers.
At the time of this writing there are more than 300 active
users of \pkg{RProtoBuf} using it to read data from and otherwise interact
with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
other languages.

The \pkg{RProtoBuf} package provides users with the ability to generate,
parse, and manipulate Protocol Buffer messages in \proglang{R}.  It is our hope that this
package will make Protocol Buffers more accessible to the \proglang{R} community, and
thereby makes a small contribution towards better integration between \proglang{R} and
other software systems and applications.

>>>>

Table~\ref{table-get-types} details the correspondence between the
field type and the type of data that is retrieved by \verb|$| and \verb|[[|
extractors.  Three types in particular need further attention due to
specific differences in the \proglang{R} language: booleans, unsigned
integers, and 64-bit integers.

@@

\Sexpr{m} data sets can be converted to Protocol Buffers
without loss of information (\Sexpr{format(100*m/n,digits=1)}\%). Upon closer
inspection, all other datasets are objects of class \texttt{nfnGroupedData}.
This class represents a special type of data frame that has some additional 
attributes (such as a \emph{formula} object) used by the \pkg{nlme} package.
Because formulas are \proglang{R} \emph{language} objects, they have little meaning to
other systems, and are not supported by the \texttt{rexp.proto} descriptor.
When \texttt{serialize\_pb} is used on objects of this class, it will serialize
the data frame and all attributes, except for the formula.

@@

\section{Summary}  % DE Simpler title
\label{sec:summary}
Over the past decade, many formats for interoperable
data exchange have become available, each with their unique features,
strengths and weaknesses.  
Text based formats such as CSV and JSON are easy to use, and will likely 
remain popular among statisticians for many years to come. However, in the 
context of increasingly complex analysis stacks and applications involving 
distributed computing as well as mixed language analysis pipelines, choosing a more 
sophisticated data interchange format may reap considerable benefits. 
The Protocol Buffers protocol and library offers a unique combination of features, performance,
maturity, and forward-compatibility that seems particulary well suited for data-driven 
applications and numerical computing.

%% DE Re-ordering so that we end on RProtoBuf
The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
extends the \proglang{R} system with the ability to create, read,
write, parse, and manipulate Protocol
Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
for the past three years by statisticians, analysts, and software engineers.
At the time of this writing there are more than 300 active
users of \pkg{RProtoBuf} using it to read data from and otherwise interact
with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
other languages.  It is our hope that this
package will make Protocol Buffers more accessible to the \proglang{R} community, and
thereby makes a small contribution towards better integration between \proglang{R} and
other software systems and applications.
